---
title: "Homework 6 Solution"
author: "Mohsen Nabian"
date: "6/25/2015"
output: html_document
---


++age++      18-29         30-44       45-59        60+        Sum
Democrat      86            72          73          71         302
INdependent   52            51          55          54         212
Republican    61            74          70          73         278
Sum           199           197         198         198        792


1.
a. Based on the exit poll results, is age independent of Party ID or not? Conduct a chi-squared test by
hand, showing each step in readably-formatted latex.
b. Verify your results using R to conduct the test.




If we have two groups each group spans the whole population, in order to prove the independency of groups, we need to have a table of samples like above as well as doing the Chi-Square analysis. Here are the hypothesis:
H0: the age and political view are INDEPENDENT.
H1: the age and political view are NOT totally independent.

Assuming null hypothesis, we can find expected value for each elemet of the sample table.
Lest calculate the number of participants in our sample. 
$$ n= (86+52+61)+(72+51+74)+(73+55+70)+(71+54+73)=792 $$
Finding expected values:
 $$ fe=\frac{(row total)\times (column total)}{overall total}$$
 
so here are the expected values table:


++age+++++    18-29         30-44       45-59        60+        
Democrat      75.88        75.1        75.5        75.5       
INdependent   53.26        52.73       53          53        
Republican    69.85        69.15       69.5        69.5  

Calculating   
$$ X^2=\sum{\frac{(fo-fe)^2}{fe}}$$

Thus
$$ X^2= \frac{(86-75.88)^2}{75.88}+\frac{(72-75.1)^2}{75.1}+\frac{(73-75.5)^2}{75.5}+\frac{(71-75.5)^2}{75.5}+\frac{(52-53.26)^2}{53.26}+\frac{(51-52.73)^2}{52.73}+\frac{(55-53)^2}{53}+\frac{(54-53)^2}{53}+\frac{(61-69.85)^2}{69.85}+\frac{(74-69.15)^2}{69.15}+\frac{(70-69.5)^2}{69.5}+\frac{(73-69.5)^2}{69.5}$$

$$x^2=1.35+0.13+0.083+0.27+0.029+0.057+0.075+0.02+1.12+0.34+0.036+0.18=3.69$$

Degree of freedom
$$df=(r-1)(c-1)=2\times 3=6$$



```{r}
pchisq(3.69,6,lower.tail=FALSE)
qchisq(0.95,6,lower.tail=FALSE)

```
So both computations above shows that the null hypothesis could be true and unrejetable. 
Lets do the whole proceadure by R function :
```{r}
ageparty<-data.frame(age1=c(86,52,61),age2=c(72,51,74),age3=c(73,55,70),age4=c(71,54,73),row.names=c("democrats","independents","republicans"))
chisq.test(ageparty)

```
The R Test also supports our hand calculations.

2.
a. Now test for independence using ANOVA (an F test). Your three groups are Democrats, Independents,
and Republicans. The average age for a Democrat is 43.3, for an Independent it's 44.6, and for a
Republican it's 45.1. The standard deviations of each are D: 9.1, I: 9.2, R: 9.2. The overall mean age is
44.2. Do the F test by hand, again showing each step.

b. Check your results in R using simulated data. Generate a simulated dataset by creating three vectors:
Democrats, Republicans, and Independents. Each vector should be a list of ages, each with a length equal
to the number of Democrats, Independents, and Republicans in the table above, and the appropriate
mean and sd based on 2.a (use rnorm to generate the vectors). Combine all three into a single dataframe
with two variables: age, and a factor that specifies D, I, or R. Then conduct an F test using R's aov
function on that data and compare the results to 2.a. Note that your results may not exactly match 2a
either quantitatively or qualitatively.


Solutions:
===================================================================================
Here is the summary of the F-Test analysis

N total samples in G different groups of one same parameter x :
$$(\bar{x_1},s_1,n_1),(\bar{x_1},s_1,n_1),...,(\bar{x_g},s_g,n_g) $$                                                              Null Hypothesis:   
$$H_0 = \mu_1=\mu_2=...=\mu_g=\bar{x}(average\,of\,all\,samples)$$  
(That means that the paramter x in all those differnt group are the same, no dependency in x to any group.)
$$H_a = at\,least\,one\,group\,has\,different\,response\,to\,x.$$
(That means There could be some dependency between the parameter and anyone of the goups.)                   
$$F-statistic = \frac{average\,variance\,between\,groups}{average\,variance\,within\,groups}$$     
N=total number of all data
G=the number of groups                  
$$average\,variance\,between\,groups=\frac{n_1(\bar{x_1}-\bar{x})^2+...+n_G(\bar{x_G}-\bar{x})^2}{G-1}$$                      $$average\,variance\,within\,groups=\frac{(n_1-1)\times (s_1^2)+...+(n_1-1)\times (s_1^2)}{N-G}$$    
degree of freedoms:
   $$df1=G-1$$      $$df2=N-G$$ 

========================================================================================

Democrates(mean=43.3,s=9.1,n=302) 
Independent(mean=44.6,s=9.2,n=212)
Republicans(mean=45.1,s=9.2,n=278)
Overal Mean=44.2
N=792, G=3

$$ average\,variance\,between\,groups=\frac{302\times(43.3-44.2)^2+212\times(44-44.2)^2+278\times(45.1-44.2)^2 }{3-1}=239.14 $$
$$ average\,variance\,within\,groups =\frac{(302-1)\times (9.1^2)+(212-1)\times (9.2^2)+(278-1)\times(9.2^2)}{792-3}= 83.94$$    
$$F-statistic=\frac{239.14}{83.94}=2.85$$
$$df1=3-1=2$$
$$df2=792-3=789$$ 
   
```{r}
pf(2.85,2,789)
qf(0.95,2,789)
```
Assuming 0.05 error, still we are UNABLE to reject the null hypothesis (0.94<0.95).But it is very close. That mean we can say with the error of 0.07 the null hypothesis is rejected. 
Lets do the F-Test:
```{r}
Democrates<-rnorm(302,43.3,901)
Independents<-rnorm(212,44.6,9.2)
Republicans<-rnorm(278,45.1,9.2)
DD<-cbind("D",Democrates)
II<-cbind("I",Independents)
RR<-cbind("R",Republicans)
age_party<-rbind(DD,II,RR)
age_party<-cbind(as.factor(age_party[,1]),as.numeric(age_party[,2]))
DF<-data.frame(age_party)
head(DF)
aov.ex<-aov(DF[,2]~DF[,1],data=DF)
summary(aov.ex)
```
Results will change significantly any time we run this code. However, it shows how we can determine dependency of age with party inclination with statistical purposes. We assume the Null hypotheis that there is no dependency and try to reject it.   





