---
title: "Homework 10 Solution"
author: "Mohsen Nabian"
date: "8/1/2015"
output: pdf_document
---

1. Using the anes_2008tr.csv dataset in Course Resources, model vote-rep (whether the respondent voted Republican in the last election) as a function of age, race, income, and ideology.

```{r}
anes_2008tr <- read.csv("C:/Users/monabiyan/SkyDrive/Summer 2015/Statistics/HW8/anes_2008tr.csv",sep=",",header=TRUE,stringsAsFactors=FALSE)

```



a. What's the probability of voting Republican for a white person of average age, income, and ideology?

```{r}
age_mean=mean(anes_2008tr$age)
income_mean=mean(anes_2008tr$income)
ideology_mean=mean(anes_2008tr$ideology_con)
RaceWhite<-1
lr1 <- glm(vote_rep ~ age + income + ideology_con+race_white,data=anes_2008tr,family="binomial")

summary(lr1)
# The nice thing about glm is that it automotically delete the rows with NA values. Here. 783 rows were not counted as mentioned in the summary.
B0=as.numeric((lr1$coef)[1])
B1=as.numeric((lr1$coef)[2])
B2=as.numeric((lr1$coef)[3])
B3=as.numeric((lr1$coef)[4])
B4=as.numeric((lr1$coef)[5])
H=exp(B0+B1*age_mean+B2*income_mean+B3*ideology_mean+B4*RaceWhite)
prob_vote_rep_white<-(H)/(1+H)
prob_vote_rep_white
```

b. What's the change in probability of voting Republican for a person of average age, income, and ideology
who switches from black to white?
```{r}
RaceBlack=0

RaceWhite=1

odd_ratio_white=exp(B0+B1*age_mean+B2*income_mean+B3*ideology_mean+B4*RaceWhite)

odd_ratio_black=exp(B0+B1*age_mean+B2*income_mean+B3*ideology_mean+B4*RaceBlack)

prob_vote_rep_white<-(odd_ratio_white)/(1+odd_ratio_white)

prob_vote_rep_black<-(odd_ratio_black)/(1+odd_ratio_black)

diff<-prob_vote_rep_white-prob_vote_rep_black

diff

```


c. Using the e formula from the lesson, what's the effect on the odds ratio of shifting from black to white?
```{r}
diff<-odd_ratio_white-odd_ratio_black
diff
```

d. What has a greater effect on voting Republican: an age increase of 50 years, or an incease of one income
bracket?

lets caclculate based on the odd ratio which make the calculations simpler. As odd ratio increases, The P(y=1) increases. 

one unit increase in x_i is equivalent to multiplying the odd with exp(Bi),two units increase in xi is equivalent to multiplying the odd with exp(2*Bi)... So
```{r}
age_increase_factor<-exp(50*B1)
age_increase_factor
income_increase_factor<-exp(1*B2)
income_increase_factor

```
So one unit increase in income has higher effect that 50 years of age increase in the probability of a person vote for a republican


e. Now run the regression with all the other variables in anes_2008tr (except for voted). How do your
coefficients change? What do you think explains any coefficient that became or lost significance?
```{r}

reg <- glm(vote_rep ~ age + income + ideology_con+ race_white+ gender_male+partyid_rep,             data=anes_2008tr,family="binomial")
summary(lr1)
summary(reg)
```
In the full regression, 'age' regains the significance, and gender_male is not significant. That means gender in fact has no effect on the vote. 'income','ideology' and 'race' still have their significance but have lower coefficients. Party-id has also significant effect. 
'age' and 'income' got less significant. 'income' could be a function of other variables like, age and race and that is why it loses its significance. 




2. Construct a simulated y variable with 100 observations where each observation (year) is a function of
the previous observation: specifically, yt is 80% yt???1 + 20% random noise with mean 0 and sd 1 (and
y1 = 1). Estimate an ARIMA model using auto.arima() from the forecast package and interpret
the results, in particular the ARIMA(p,d,q) numbers and the coefficients reported, if any. What do you
think is going on here?

```{r}

library(forecast)
yy=0
yy[1]=1
for (i in 2:100)
{
 yy[i]<-0.8*yy[i-1]+0.2*rnorm(n=1,mean=0,sd=1)
}
auto.arima(yy)

```

The auto arima sayas AR=1, Differencing=0, MA=0
that means y(t) is only dependent to y(t-1) and not more lags. which corresponds to our expectations.
Moreover, 0 differencing is that y is not a time series variable which also make sense since there is no time dependency for y.

Finally 0 for MA means there is no moving average. That is true beacuase every element depends on the preivious element as well as external noises. 

The results also suggests 0.8627 for the coefficient for y(t-1) which is close to 0.8 that we chose in our formula.

3. Find some existing data that either has temporal or bindary dependent variable data and run a ARIMA
or logit model on it and interpret the results in detail. You can use data from the previous assignment,
and you can construct a binary variable out of some existing continuous variable if you like.


```{r}
mydata <- read.csv("http://www.ats.ucla.edu/stat/data/binary.csv")
## view the first few rows of the data
head(mydata)
admission_reg<- glm(admit ~ gre + gpa + rank ,data=mydata ,family="binomial")
summary(admission_reg)

```


accordig to the result, 'rank' plays an important role on acceptance of the applicants.Of course lower rank means more qualified applicant and thats why we see negetive coefficient.
However, GRE and GPA are important but less significant.1 reason that Gre coefficient is very low is that the gre scale is 100-800 and much higher to rank and gpa. 


