---
title: "Homework 1 Solution"
author: "Mohsen Nabian"
date: "5/19/2015"
output: html_document
---

1. 

a. Create two vectors named v1 and v2, where v1 is the sequence of integers from 2 to 6, and v2 is the sequence of integers from 5 to 9.  

```{r}
v1=c(2:6)
v1
v2=c(5:9)
v2

productv=v1%*%v2
productv

```
b.What is v2 minus v1?
```{r}
minusv=v2-v1
minusv
```

c.What is the inner product of v1 and v2?
```{r}
productv=v1%*%v2
productv
```
d.Replace the elements in v1+v2 that are greater than 10 with the number 0. Show that vector.
```{r}
sumv=v1+v2
sumv[which(sumv>10)]=0
sumv
```

2. 

a. Create a 5 by 5 matrix with the numbers 1 to 25 as its elements, and call it m1.

```{r}
m1=matrix(1:25,nrow=5)
m1

```

b. What is m1 times v1?

```{r}
m1v1=m1%*%v1
m1v1
```

c.What is v1 times m1?
```{r}
v1m1=v1%*%m1
v1m1
```

d.What is m1 times the transpose of m1?
```{r}
m1trm1=m1%*%t(m1)
m1trm1
```

3.Create a date frame with at least five rows and three columns. The first variable (column) should be dates, the second variable should be strings (characters), and the third variable should be numbers. Name each variable something appropriate and short. 


```{r}
v1=c(2:6)
v2=c(5:9)
m1=matrix(1:25,nrow=5)
getwd()
setwd("C:/Users/nabian.m/OneDrive/Summer 2015/Statistics/HW1")
getwd()
df1<-data.frame(c("2015-05-19","2015-05-19","2015-05-19","2015-05-19","2015-05-19"),c("Mason","Kati","Sara","David","Jack"),c(80,90,95,100,50))
colnames(df1)<-c("Submission_Date","Name","Grade")
df1$Submission_Date<-as.Date(df1$Submission_Date)
df1$Name<-as.character(df1$Name)
df1$Grade=as.numeric(df1$Grade)
str(df1)
write.table(df1,file="df1.csv",row.names=FALSE,sep=",")
df1 <- read.table(file="df1.csv",header=TRUE,sep=",",stringsAsFactors=FALSE)
str(df1)
df2=df1[c(1,3,5),c(1,2)]
df2
df1[df1[1:5,3]%%2==0,3]=0
df1
list1=list(v1,v2,m1,df1)
names(list1) <- c("vector1","vector2","matrix1","data_frame")
str(list1)
list1$matrix1[2,1]
```


4. 

a. Using latex equation notation in your .Rmd file, write out the quadratic formula, so that in your html file it looks pretty and like the version we all learned in high school. (Eg, see the box in the top right of this wikipedia page: <http://en.wikipedia.org/wiki/Quadratic_equation>.)





$$\frac{-b\pm \sqrt{b^2 - 4ac}}{2a}$$
