---
title: "Homework 7 Solution"
author: "Mohsen Nabian"
date: "7/09/2015"
output: pdf_document
---

Using a high-dimensional dataset of your choice, perform a factor analysis and clustering and interpret the
results. You may use, for instance, the datasets inside the psych package, such as bfi (25 personality items
thought to boil down to a few core personality types) or iqitems (14 scores that are thought to boil down to
a few core mental skills), or anything else you can find. (Load the data using, for instance, data(bfi) after
loading the psych package; you may need to clean it a bit first with na.omit() to remove the observations
with na items, or else impute those missing items.)
For the factor analysis, you may use any of the methods covered in the lesson - they should all produce
similar results, though princomp and prcomp might be simplest. You don't have to interpret everything, say,
fa() outputs, which is a lot of stuff - easier to use str() to examine the output of your function and find
the quantities you want.


Q1)  After running your factor analysis or PCA, be sure to discuss and interpret your output:

1. Examine the factor eigenvalues or variances (or the sdev or standard deviations as reported by prcomp
or princomp, which you then need to square to get the variances). Plot these in a scree plot and use
the "elbow" test to guess how many factors one should retain. What proportion of the total variance
does your subset of variables explain?
2. Examine the loadings of the factors on the variables (sometimes called the "rotation" in the function
output) - ie, the projection of the factors on the variables - focusing on just the first one or two factors.
Sort the variables by their loadings, and try to interpret what the first one or two factors "mean." This
may require looking more carefully into the dataset to understand exactly what each of the variables
were measuring. You can find more about the data in the psych package using ?psych or visiting
http://personality-project.org/ .


Solution:

Factor analysis proceadure:

1)Standardize the variables
2)Create the covariance matrix
3)Find the eigenvectors and eigenvalues for that matrix.
4)Choose how many of them we want to keep and analyze.

```{r}

library(psych)  #have some data as well as some machine learning functions like fa()
data(bfi)
data(bfi.dictionary)
bfi.dictionary
bfi_data_not_scaled <- na.omit(bfi) 
varnames<-names(bfi_data_not_scaled)    #keeping the names to use later
```
But we need to linearly scale the data into the same range. To do this, I write the following function.

```{r}  
linMap <- function(DF, from, to)   #linear mapping
{
 
  for (i in 1:ncol(DF))
  {
    x<-DF[,i]
    DF[,i]<-((x - min(x)) / max(x - min(x))) * (to - from) + from
    
  }
 
  return(DF)
}
```

I will scale data linearly in [0,100]

```{r}
bfi_data<-linMap(bfi_data_not_scaled,0,100)  #normalized in the scale of -100, 100

names(bfi_data)<-varnames
dim(bfi_data)
head(bfi_data)
tail(bfi_data)
```

now factor analysis:

```{r}
eigenm <- eigen(cov(bfi_data))      # Calculating Eigen Values and Eigne Vectors
varnames<-names(bfi_data)

################ First Factor
eigen1 <- eigenm$vectors[,1]
factor1<-data.frame(varnames[order(eigen1)],eigen1[order(eigen1)]) # making a data frame putting data in the order of coefficients magnitude.
names(factor1)<-c("variable","coeff")
head(factor1)
################ Second Factor
eigen2 <- eigenm$vectors[,2]
factor2<-data.frame(varnames[order(eigen2)],eigen2[order(eigen2)])
names(factor2)<-c("variable","coeff")
head(factor2)
################ Third Factor
eigen3 <- eigenm$vectors[,3]
factor3<-data.frame(varnames[order(eigen3)],eigen3[order(eigen3)])
names(factor3)<-c("variable","coeff")
head(factor3)
################ Forth Factor
eigen4 <- eigenm$vectors[,4]
factor4<-data.frame(varnames[order(eigen4)],eigen4[order(eigen4)])
names(factor4)<-c("variable","coeff")
head(factor4)
#####################
################ 5th Factor
eigen5 <- eigenm$vectors[,5]
factor5<-data.frame(varnames[order(eigen5)],eigen5[order(eigen5)])
names(factor5)<-c("variable","coeff")
head(factor5)
################ 6th Factor
eigen6 <- eigenm$vectors[,6]
factor6<-data.frame(varnames[order(eigen6)],eigen6[order(eigen6)])
names(factor6)<-c("variable","coeff")
head(factor6)
################ 7th Factor
eigen7 <- eigenm$vectors[,7]
factor7<-data.frame(varnames[order(eigen7)],eigen4[order(eigen7)])
names(factor7)<-c("variable","coeff")
head(factor7)
#####################
############Eigen Values
eigen_value<-eigenm$values
plot(eigen_value,type="b")
```
So based on the "Elbow" rule, we would pick 7 factors as our prinipals and assume the rest as noises.

Looking further in factor 1: 
```{r}
factor1
plot(factor1)
```

Factor 1: High in E4 and E2
That means peoples personality is dominantly dependent on the factor that wether they are socail and make friends or they are not good at socializing.


```{r}
factor2
plot(factor2)
```
Factor 2: 
Gender in this facor plays a dominant role. That means gender could paly a deteministic role in people's personality characterizations. 
```{r}
factor3
plot(factor3)
```
Factor3: 
In this factor also gender plays the dominant role.

Summary: In this analysis we might say peoples personality Questionnaire highly revolves around the gender and social skills questions and these are key questions. 


########################################

Q2)   Next perform a cluster analysis of the same data.
3. First use k-means and examine the centers of the first two or three clusters. How are they similar to
and different from the factor loadings of the first couple factors?
4. Next use hierarchical clustering. Print the dendrogram, and use that to guide your choice of the number
of clusters. Use cutree to generate a list of which clusters each observation belongs to. Aggregate
the data by cluster and then examine those centers (the aggregate means) as you did in (3). Can you
interpret all of them meaningfully using the methods from (3) to look at the centers?
5. From the factor and cluster analysis, what can you say more generally about what you have learned
about your data?




```{r}
hout <- hclust(dist(bfi_data),method="complete") #or method="average" or...
plot(hout,labels=FALSE)

#To get the cluster assignments, we just apply cutree to the hclust output,
#choose either the height to cut it at, or the number of clusters:
#as.vector(cutree(hout,h=420)) # cut the plot at height=21
abline(a=420,b=0,col="red")

```

According to the plot,having 3 clusters seems to be reasonable. Lets do the K-means with 3 clusters:

```{r}
set.seed(100)

kout <- kmeans(bfi_data,centers=3,nstart=30)  #means 30 times change the random initialization of the clustring
                                              # and choose the best one that has minimum                                                      average distance

centroids <- kout$centers

topvars_centroid1 <- centroids[1,order(centroids[1,])]
topvars_centroid2 <- centroids[2,order(centroids[2,])]
topvars_centroid3 <- centroids[3,order(centroids[3,])]



topvars_centroid1
topvars_centroid2
topvars_centroid3

```

cluster 1 and 3 are specified for men. and cluster 2 is for women. 
according to cluster 1, A4, A2 A5 and E4,A3 are high for men.
That means these men are:
  1)loving children
  2)Inquire about others' well-being
  3)Make people feel at ease
  4)Make friends easily
  5)Know how to comfort others
These Men are so social and caring about others.


however, O4 and O1 are high in women.
These women are:
  1) Spending time reflecting on things
  2) full of ideas
apparently  alot of these women are passionate about their future, thinking alot and not so social.

The 3rd cluster which is again for men:

They are high in O4 and significantly less in A4,A2 and A5
These means that these men are 
  1)full of ideas
  2) Less social
These men are more serious about the life. 


Summary:

Factor analysis and Cluster analysis in this study provides the fact that people's personality could be divided majorly based on their gender and social skills.
Those having the same sex and social skills would be having almost the same other personality characteristics. 
Also  cluster analysis, and the hierarchy plot, demonstrated us that socialability in men are forming their other characteristics which is not the case in women as much. 






